\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=.75in]{geometry}




\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Now begins the stuff that I added.
% ----------------------------------

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\large\bf}
\subsectionfont{\rmfamily\mdseries\normalsize\itshape}


% Make lists without bullets
%\renewenvironment{itemize}{
%  \begin{list}{}{
%    \setlength{\leftmargin}{1.5em}
%  }
%}{
%  \end{list}
%}


% Make parskips rather than indent with lists.
\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

% Use fontawesome. Note: you'll need TeXLive 2015. Update.
\usepackage{fontawesome}

% Fancyhdr, as I tend to do with these personal documents.
\usepackage{fancyhdr,lastpage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}
\lhead{}
\chead{}
\rhead{}
\lfoot{
\cfoot{\scriptsize  Marwin Machay Indio do Brasil do Carmo - CV -  }}
\rfoot{\scriptsize \thepage/{\hypersetup{linkcolor=black}\pageref{LastPage}}}

% Always load hyperref last.
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref

\hypersetup{unicode=true,
            pdftitle={Marwin Machay Indio do Brasil do Carmo:  CV (Curriculum Vitae)},
            pdfauthor={Marwin Machay Indio do Brasil do Carmo},
            pdfkeywords={RMarkdown, academic CV, template},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true, bookmarks=true}
\urlstyle{same}  % don't use monospace font for urls

\begin{document}


\centerline{\huge \bf Marwin Machay Indio do Brasil do Carmo}

\vspace{2 mm}

\hrule

\vspace{2 mm}



\moveleft.5\hoffset\centerline{ \faEnvelopeO \hspace{1 mm} \href{mailto:}{\tt \href{mailto:mmcarmo@ucdavis.edu}{\nolinkurl{mmcarmo@ucdavis.edu}}} \hspace{1 mm}  \faGlobe \hspace{1 mm} \href{http://marwincarmo.github.io}{\tt marwincarmo.github.io} \hspace{1 mm}   \faGithub \hspace{1 mm} \href{http://github.com/marwincarmo}{\tt marwincarmo}   | \emph{Updated:} \today} 

\vspace{2 mm}

\hrule


\section{EDUCATION}\label{education}

\hrule

\textbf{University of California, Davis}, Ph.D.~Quantitative Psychology
\hfill 2023--Present\\
\emph{Advisor}: \href{https://rastlab.ucdavis.edu/}{Philippe Rast,
Ph.D.}\\
\emph{Research topics}: Intra-individual variability, Mixed-effects
location scale models, Bayesian inference, Gaussian graphical models\\
\textbf{University of São Paulo}, M.Sc. Psychiatry \hfill 2021--2023\\
Recipient of a Master's Fellowship from Sao Paulo Research Foundation
(FAPESP)\\
\emph{Advisor}:
\href{https://scholar.google.com.br/citations?hl=pt-BR&user=sLWIwcoAAAAJ}{Renatha
El-Rafihi Ferreira, Ph.D.}\\
\textbf{Rio de Janeiro State University}, Bachelor in Psychology
\hfill 2014--2019\\
\textbf{Federal Institute of Education, Science and Technology}, B.Tech.
Chemical Processes \hfill 2008--2013

\section{PUBLICATIONS}\label{publications}

\hrule

\subsection{\texorpdfstring{\textbf{Peer-reviewed}}{Peer-reviewed}}\label{peer-reviewed}

* \emph{denotes manuscript in Portuguese}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Rafihi-Ferreira, R., Hasan, R., Toscanini, A. C., Linares, I. M. P.,
  Borges, D. S., Brasil, I. P., \textbf{Carmo, M.}, Neto, F. L., Morin,
  C. (2024). Acceptance and commitment therapy versus cognitive
  behavioral therapy for insomnia: A randomized controlled trial.
  \emph{Journal of Consulting and Clinical Psychology}.
  \url{https://doi.org/10.1037/ccp0000881}
\item
  \textbf{Carmo, M. M. I. B.}, Leme, V. B. R., Quiterio, P. L. (2024).
  Social skills and people with disabilities: overview of Brazilian
  research. \emph{Psicologia Escolar e Educacional}.
  \url{https://doi.org/10.1590/2175-35392024-246235-T}
\item
  Rafihi-Ferreira, R., \textbf{Carmo, M. M. I. B.}, Bassoli, L., Hasan,
  R., Aizawa, I. M. I., \& Toscanini, A. C. (2022). Cognitive and
  psychological factors associated with severe insomnia in Brazilian
  women: A cross-sectional study. \emph{Psicologia: Reflexão e Crítica}.
  \url{https://doi.org/10.1186/s41155-022-00243-x}
\item
  *Quiterio, P. L., Nunes, L. R. O. P., Camelo, B. L., Silva, J. P., \&
  \textbf{Carmo, M. M. I. B.} (2021). Promoting the social skills of
  future teachers focusing on inclusive education. \emph{Estudos e
  Pesquisa em Psicologia}. \url{https://doi.org/h43p}
\item
  *Quiterio, P. L., Leme, V. B. R., \textbf{Carmo, M. M. I. B.}, Silva,
  J. P., \& Camelo, B. L. (2021). Family guidance program developed in
  the applied psychology service of a public university.
  \emph{Psicologia Clínica}.
\item
  *\textbf{Carmo, M. M. I. B.}, Gilla, C. G., \& Quiterio, P. L. (2020).
  A study about the inclusion of people with disabilities in the
  Brazilian labour market. \emph{Interação em Psicologia}.
  \url{https://doi.org/h43q}
\item
  Ferreira, W. H., \textbf{Carmo, M. M. I. B.}, Silva, A. L. N., \&
  Andrade, C. T. (2014). Effect of structure and viscosity of the
  components on some properties of starch-rich hybrid blends.
  \emph{Carbohydrate polymers}. \url{https://doi.org/f6w475}
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Under
Review}}{Under Review}}\label{under-review}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Carmo, M.}, Williams, D. R., \& Rast, P. Beyond Average
  Scores: Identification of Consistent and Inconsistent Academic
  Achievement in Grouping Units. Under review at \emph{Journal of
  Educational and Behavioral Statistics}.
  {[}\href{https://doi.org/10.31234/osf.io/sh6ne}{preprint}{]}
\item
  \textbf{Carmo, M.}, Rafihi-Ferreira, R., Paulos-Guarnieri, L., \&
  Pires, M. L. N. The Sleep Acceptance Scale (SAS): Development,
  Psychometric Properties, and Applications. Under review at
  \emph{Behavioral Sleep Medicine.}.
\item
  Conway, B. A., \textbf{Carmo, M. M. I. B.}, Filho, H. S. L. S., Alves,
  M. M., Hasan, R., Toscanini, A. C., \& Rafihi-Ferreira, R. Personality
  traits and insomnia: direct and anxiety-mediated associations. Under
  review at \emph{Journal of Sleep Research}.
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Working
Papers}}{Working Papers}}\label{working-papers}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Carmo, M.}, Morin, C. M., Hasan, R., Toscanini, A. C., \&
  Rafihi-Ferreira, R. \emph{The network of dysfunctional beliefs about
  sleep: a structural re-analysis of the DBAS-16}.
  {[}\href{https://osf.io/preprints/psyarxiv/6vuf3}{preprint}{]}
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Blogs}}{Blogs}}\label{blogs}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Carmo, M. M. I. B.} (2022). Simulating post selection
  inference.
  {[}\href{https://marwincarmo.github.io/posts/2021-12-20-model-selection-bias/}{post}{]}
\item
  \textbf{Carmo, M. M. I. B.} (2021). Regression Artifact - Re-running
  Farmus et. al (2019).
  {[}\href{https://marwincarmo.github.io/posts/2021-11-04-simulating-regression-artifact/}{post}{]}
\end{enumerate}

\section{SOFTWARE DEVELOPMENT}\label{software-development}

\hrule

\textbf{\texttt{ivd}}. An R package for random effects selection in the
scale part of Mixed Effects Location Scale Models using a spike-and-slab
prior on the random effects of the scale.
{[}\href{https://github.com/consistentlyBetter/ivd}{Github repo}

\textbf{Model selection app}. A Shiny app designed for performing
simulations of statistical inferences after model selection procedures.
Its purpose is to illustrate the problems that arise when model
selection, parameters estimation and statistical inferences are
undertaken with the same data set.
{[}\href{https://marwin.shinyapps.io/model-selection-app/}{App}{]}
{[}\href{https://github.com/marwincarmo/model-selection-app}{Github
repo}{]}

\textbf{Brazilian Research Funding app}. In this app I explore the open
database of the National Council for Scientific and Technological
Development (CNPq), a Brazilian Government research funding agency.
{[}\href{https://marwin.shinyapps.io/bolsas-cnpq/}{App}{]}
{[}\href{https://github.com/marwincarmo/shiny-bolsas}{Github repo}{]}

\section{RESEARCH AND WORK
EXPERIENCE}\label{research-and-work-experience}

\hrule

\textbf{University of California, Davis} \hfill September
2023--Present\\
\emph{Graduate Student Teaching Assistant} \hfill Davis, CA\\
- Statistical Analysis of Psychological Data II (3 quarters)\\
- Data Visualization in the Social Sciences (1 quarter)\\
- General Psychology (1 quarter)

\textbf{Psicometria Online Academy} \hfill May 2022--October 2024\\
\emph{Virtual Teaching Assistant} (Psychometrics) \hfill Rio de Janeiro,
Brazil (\emph{remote})

\textbf{PsicoDados} \hfill January 2022--December 2022\\
\emph{Statistical Consultant} \hfill Rio de Janeiro, Brazil
(\emph{remote})

\textbf{Psychology of Human Development Laboratory - UERJ} \hfill August
2019--August 2020\\
\emph{Research Assistant} \hfill Rio de Janeiro, Brazil\\
Recipient of FAPERJ's Training and Technical Qualification fellowship

\textbf{Socio-Emotional Development, Inclusion and Social Skills
Laboratory - UERJ} \hfill July 2017--July 2019\\
\emph{Undergraduate Research Project} \hfill Rio de Janeiro, Brazil\\
Recipient of UERJ's Department of Scholarships funding for undergraduate
research project

\textbf{Biodegradable Polymer Research Laboratory - UFRJ} \hfill June
2011--March 2013\\
\emph{Undergraduate Research Project} \hfill Rio de Janeiro, Brazil

\section{OTHER RELEVANT EXPERIENCE}\label{other-relevant-experience}

\hrule

\textbf{\href{https://measuringpsychology.netlify.app/}{Measuring
Psychology Study Goup and Community}} \hfill February 2022--October
2023\\
\emph{Founder and facilitator} \hfill Online

\begin{itemize}
\tightlist
\item
  Created a global community of 200+ members interested in discussions,
  collaborations, and networking in the field of measurement in the
  human sciences;
\item
  Organized bi-weekly meetings and led study sessions on measurement
  theory;
\item
  Orchestrated invited presentations by renowned scholars like Dr.~Derek
  C. Briggs, Dr.~Laura F. Bringmann, Dr.~Markus I. Eronen, Dr.~Alexander
  Christensen, Dr.~James Grice, and others.
\end{itemize}

\section{SKILLS}\label{skills}

\hrule

\emph{Programming}

\begin{itemize}
\tightlist
\item
  R: exploratory data analysis, data cleaning, data visualization,
  statistical data analyses, programmatic reporting (R Markdown/
  Quarto), web scraping, dashboards, packages development, Shiny apps
  development.
\item
  Python: exploratory data analysis, data cleaning, data visualization.
\end{itemize}

\emph{Tools}

\begin{itemize}
\tightlist
\item
  Statistical softwares: R, SPSS, JASP, JAMOVI
\item
  Reporting: RMarkdown, Quarto, \LaTeX
\item
  Version control: Git and Github
\item
  Reference management: Zotero
\item
  Data collection and management: REDCap
\end{itemize}

\section{RELEVANT COURSEWORK}\label{relevant-coursework}

\hrule

\emph{Mathematics \& Statistics}\\
Multivariable Calculus, Linear Algebra, Probability Theory, Machine
Learning, Structural Equation Modeling, Applied Bayesian Statistics.

\emph{Programming}\\
Monte Carlo Simulations, Data Visualization, Data Cleaning and
Management.

\section{FIRST-AUTHOR PRESENTATIONS}\label{first-author-presentations}

\hrule

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  2024 -- Poster presentation: A holistic view of academic performance:
  beyond averages with MELSM and spike- and-slab. \emph{89th Annual
  Meeting of the Psychometric Society}, Prague, Czech Republic.
  {[}\href{https://marwincarmo.github.io/posters/imps24}{poster}{]}
\item
  2022 -- Poster presentation: Development of the Sleep Acceptance Scale
  (SAS). \emph{XIX Brazilian Congress of Sleep}, Goiania, Brazil.
\item
  2022 -- Oral presentation: Understanding model selection bias via
  simulation. \emph{VI International Seminar on Statistics with R},
  Online.
  {[}\href{https://marwincarmo.github.io/slides/vi-ser/selection-bias}{slides}{]}{[}\href{https://www.youtube.com/watch?v=eyd1VCAX864&t=910s}{video}{]}
\item
  2022 -- Oral presentation: A network approach of Insomnia and
  Dysfunctional Beliefs and Attitudes About Sleep. \emph{VII Psychiatric
  Clinic Congress}, São Paulo, Brazil. \textbf{Recipient of Best Oral
  Presentation Award}
  {[}\href{https://marwincarmo.github.io/slides/viiccp/dbas-network_slides}{slides}{]}
\item
  2022 -- Poster presentation: Network structure of acceptance of sleep
  difficulties and insomnia. \emph{VII Psychiatric Clinic Congress}, São
  Paulo, Brazil.
  {[}\href{https://marwincarmo.github.io/posters/poster_viicp_spaq}{poster}{]}
\item
  2021 -- Poster presentation: The association between sleep-related
  cognitions, psychological flexibility, and insomnia symptoms.
  \emph{XVIII Brazilian Congress of Sleep}, São Paulo, Brazil.
  \textbf{Recipient of Top Posters Award}
  {[}\href{https://marwincarmo.github.io/posters/cb_sono_poster}{poster}{]}
  {[}\href{https://sono2021.com.br/evento/sono2021/trabalhosaprovados/naintegra/1047}{abstract}{]}
\item
  2019 -- Poster presentation: Effects of a parent training program to
  clinical and non-clinical samples. \emph{VII International Seminar on
  Social Skills}, Maranhão, Brazil. \textbf{Recipient of 3rd place for
  Best Poster Award}
\item
  2018 -- Poster presentation: A study on the inclusion of people with
  disabilities in the Brazilian labor market. \emph{12th Regional
  Exhibition of Practices in Psychology}, Rio de Janeiro, Brazil.
\item
  2018 -- Poster presentation: Assessment of the effects of a parent
  training program. \emph{VII Brazilian Congress of Cognitive Therapy
  for Children and Adolescents}, Bahia, Brazil. \textbf{Recipient of 3rd
  place for Best Poster Award}
\end{enumerate}

\end{document}