\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=.75in]{geometry}




\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Now begins the stuff that I added.
% ----------------------------------

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\large\bf}
\subsectionfont{\rmfamily\mdseries\normalsize\itshape}


% Make lists without bullets
%\renewenvironment{itemize}{
%  \begin{list}{}{
%    \setlength{\leftmargin}{1.5em}
%  }
%}{
%  \end{list}
%}


% Make parskips rather than indent with lists.
\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

% Use fontawesome. Note: you'll need TeXLive 2015. Update.
\usepackage{fontawesome}

% Fancyhdr, as I tend to do with these personal documents.
\usepackage{fancyhdr,lastpage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}
\lhead{}
\chead{}
\rhead{}
\lfoot{
\cfoot{\scriptsize  Marwin Machay Indio do Brasil do Carmo - CV -  }}
\rfoot{\scriptsize \thepage/{\hypersetup{linkcolor=black}\pageref{LastPage}}}

% Always load hyperref last.
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref

\hypersetup{unicode=true,
            pdftitle={Marwin Machay Indio do Brasil do Carmo:  CV (Curriculum Vitae)},
            pdfauthor={Marwin Machay Indio do Brasil do Carmo},
            pdfkeywords={RMarkdown, academic CV, template},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true, bookmarks=true}
\urlstyle{same}  % don't use monospace font for urls

\begin{document}


\centerline{\huge \bf Marwin Machay Indio do Brasil do Carmo}

\vspace{2 mm}

\hrule

\vspace{2 mm}



\moveleft.5\hoffset\centerline{ \faEnvelopeO \hspace{1 mm} \href{mailto:}{\tt \href{mailto:mmcarmo@ucdavis.edu}{\nolinkurl{mmcarmo@ucdavis.edu}}} \hspace{1 mm}  \faGlobe \hspace{1 mm} \href{http://marwincarmo.github.io}{\tt marwincarmo.github.io} \hspace{1 mm}   \faGithub \hspace{1 mm} \href{http://github.com/marwincarmo}{\tt marwincarmo}   | \emph{Updated:} \today} 

\vspace{2 mm}

\hrule


\section{EDUCATION}\label{education}

\hrule

\textbf{University of California, Davis}, Ph.D.~Quantitative Psychology
\hfill 2023--Present\\
\emph{Advisor}: \href{https://rastlab.ucdavis.edu/}{Philippe Rast,
Ph.D.}\\
\emph{Research topics}: Intra-individual variability, Mixed-effects
location scale models, Bayesian inference, Gaussian graphical models\\
\textbf{University of São Paulo}, M.Sc. Psychiatry \hfill 2021--2023\\
Recipient of a Master's Fellowship from Sao Paulo Research Foundation
(FAPESP)\\
\emph{Advisor}:
\href{https://scholar.google.com.br/citations?hl=pt-BR&user=sLWIwcoAAAAJ}{Renatha
El-Rafihi Ferreira, Ph.D.}\\
\textbf{Rio de Janeiro State University}, Bachelor in Psychology
\hfill 2014--2019

\section{PUBLICATIONS}\label{publications}

\hrule

\subsection{\texorpdfstring{\textbf{Peer-Reviewed Journal
Articles}}{Peer-Reviewed Journal Articles}}\label{peer-reviewed-journal-articles}

* \emph{denotes manuscript in Portuguese}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Rafihi-Ferreira, R., Hasan, R., Toscanini, A. C., Linares, I. M. P.,
  Borges, D. S., Brasil, I. P., \textbf{Carmo, M.}, Neto, F. L., Morin,
  C. (2024). Acceptance and commitment therapy versus cognitive
  behavioral therapy for insomnia: A randomized controlled trial.
  \emph{Journal of Consulting and Clinical Psychology}.
  \url{https://doi.org/10.1037/ccp0000881}
\item
  \textbf{Carmo, M. M. I. B.}, Leme, V. B. R., Quiterio, P. L. (2024).
  Social skills and people with disabilities: overview of Brazilian
  research. \emph{Psicologia Escolar e Educacional}.
  \url{https://doi.org/10.1590/2175-35392024-246235-T}
\item
  Rafihi-Ferreira, R., \textbf{Carmo, M. M. I. B.}, Bassoli, L., Hasan,
  R., Aizawa, I. M. I., \& Toscanini, A. C. (2022). Cognitive and
  psychological factors associated with severe insomnia in Brazilian
  women: A cross-sectional study. \emph{Psicologia: Reflexão e Crítica}.
  \url{https://doi.org/10.1186/s41155-022-00243-x}
\item
  *Quiterio, P. L., Nunes, L. R. O. P., Camelo, B. L., Silva, J. P., \&
  \textbf{Carmo, M. M. I. B.} (2021). Promoting the social skills of
  future teachers focusing on inclusive education. \emph{Estudos e
  Pesquisa em Psicologia}. \url{https://doi.org/h43p}
\item
  *Quiterio, P. L., Leme, V. B. R., \textbf{Carmo, M. M. I. B.}, Silva,
  J. P., \& Camelo, B. L. (2021). Family guidance program developed in
  the applied psychology service of a public university.
  \emph{Psicologia Clínica}.
\item
  *\textbf{Carmo, M. M. I. B.}, Gilla, C. G., \& Quiterio, P. L. (2020).
  A study about the inclusion of people with disabilities in the
  Brazilian labour market. \emph{Interação em Psicologia}.
  \url{https://doi.org/h43q}
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Under
Review}}{Under Review}}\label{under-review}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Carmo, M.}, Williams, D. R., \& Rast, P. Beyond Average
  Scores: Identification of Consistent and Inconsistent Academic
  Achievement in Grouping Units. Under review at \emph{Journal of
  Educational and Behavioral Statistics}.
  {[}\href{https://doi.org/10.31234/osf.io/sh6ne}{preprint}{]}
\item
  \textbf{Carmo, M.}, Rafihi-Ferreira, R., Paulos-Guarnieri, L., \&
  Pires, M. L. N. The Sleep Acceptance Scale (SAS): Development,
  Psychometric Properties, and Applications. Under review at
  \emph{Behavioral Sleep Medicine.}.
\item
  Conway, B. A., \textbf{Carmo, M. M. I. B.}, Filho, H. S. L. S., Alves,
  M. M., Hasan, R., Toscanini, A. C., \& Rafihi-Ferreira, R. Personality
  traits and insomnia: direct and anxiety-mediated associations. Under
  review at \emph{Journal of Sleep Research}.
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Manuscripts in
Preparation}}{Manuscripts in Preparation}}\label{manuscripts-in-preparation}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Carmo, M.}, Morin, C. M., Hasan, R., Toscanini, A. C., \&
  Rafihi-Ferreira, R. \emph{The network of dysfunctional beliefs about
  sleep: a structural re-analysis of the DBAS-16}.
  {[}\href{https://osf.io/preprints/psyarxiv/6vuf3}{preprint}{]}
\item
  Ingram, E. O., \textbf{Carmo, M.}, Manderino, L., Rast, P., \& Karr,
  J. E. \emph{Exploratory Analysis of Post-Concussion-Like Symptom
  Networks in College Students with and Without a Remote History of
  Concussion}. {[}In preparation for submission{]}
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Blogs}}{Blogs}}\label{blogs}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Carmo, M. M. I. B.} (2022). Simulating post selection
  inference.
  {[}\href{https://marwincarmo.github.io/posts/2021-12-20-model-selection-bias/}{post}{]}
\item
  \textbf{Carmo, M. M. I. B.} (2021). Regression Artifact - Re-running
  Farmus et. al (2019).
  {[}\href{https://marwincarmo.github.io/posts/2021-11-04-simulating-regression-artifact/}{post}{]}
\end{enumerate}

\section{PRESENTATIONS}\label{presentations}

\hrule

\subsection{\texorpdfstring{\textbf{Selected Conference
Presentations}}{Selected Conference Presentations}}\label{selected-conference-presentations}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Carmo, M.}, Williams, D. R., \& Rast, P. (2024, July). A
  holistic view of academic performance: beyond averages with MELSM and
  spike- and-slab. \emph{89th Annual Meeting of the Psychometric
  Society}, Prague, Czech Republic.
  {[}\href{https://marwincarmo.github.io/posters/imps24}{poster}{]}
\item
  Ingram, E. O., \textbf{Carmo, M.}, Manderino, L., Rast, P., Karr, J.
  E. (2025, February). Network centrality of post-concussion symptoms in
  young adults with and without a remote history of concussion.
  \emph{Annual Rehabilitation Psychology Conference}, Louisville,
  Kentucky, USA.
\item
  Ingram, E. O., \textbf{Carmo, M.}, Manderino, L., Rast, P., Karr, J.
  E. (2025, March). A network approach to post-concussion symptoms:
  Comparing central symptoms between young adults with and without
  concussion history. \emph{World Congress on Brain Injury by the
  International Brain Injury Association}, Montreal, Canada.
\end{enumerate}

\subsection{\texorpdfstring{\textbf{Departmental
Presentations}}{Departmental Presentations}}\label{departmental-presentations}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Beyond Average Scores: Identification of Consistent and Inconsistent
  Academic Achievement in Grouping Units (Winter 2025). \emph{UC Davis
  Quantitative Research Seminar}.
\item
  A holistic view of academic performance: Beyond Averages with MELSM
  and Spike-and-Slab (Spring 2024). \emph{UC Davis Quantitative Research
  Seminar}.
\end{enumerate}

\section{METHODS ESTABLISHED}\label{methods-established}

\hrule

\textbf{Spike-and-Slab Mixed-Effects Location Scale Model (SS-MELSM)}:
The SS-MELSM is a technique for simultaneously modeling location and
scale parameters while incorporating a spike-and-slab prior to select or
shrink random effects. It involves identifying clusters with unusually
large or small within-cluster variance in academic achievement, which
can indicate inconsistent or consistent performance. This method is
implemented in the R package
\href{https://github.com/consistentlyBetter/ivd}{\texttt{ivd}}.

\section{TEACHING EXPERIENCE}\label{teaching-experience}

\hrule

\textbf{University of California, Davis} \hfill September
2023--Present\\
\emph{Graduate Student Teaching Assistant} \hfill Davis, CA

\begin{itemize}
\tightlist
\item
  PSC103B: Statistical Analysis of Psychological Data -- \textbf{3
  quarters} (Winter 24, Spring 24, and Winter 25)

  \begin{itemize}
  \tightlist
  \item
    Led laboratory sessions, graded assignments and exams, and held
    office hours.
  \end{itemize}
\item
  PSC012Y: Data Visualization in the Social Sciences -- \textbf{1
  quarter} (Fall 24)

  \begin{itemize}
  \tightlist
  \item
    Led laboratory sessions, graded assignments, and held office hours.
  \end{itemize}
\item
  PSC001: General Psychology -- \textbf{1 quarter} (Fall 23)

  \begin{itemize}
  \tightlist
  \item
    Graded assignments and held office hours.
  \end{itemize}
\end{itemize}

\textbf{Psicometria Online Academy} \hfill May 2022--October 2024\\
\emph{Virtual Teaching Assistant} (Psychometrics) \hfill Remote

\begin{itemize}
\tightlist
\item
  Provided support to students' questions in topics such as general
  linear models, multivariate statistics, structural equation modeling,
  factor analysis, item response theory and research methods.
\end{itemize}

\section{MENTORSHIP ROLES}\label{mentorship-roles}

\hrule

\textbf{Undergraduate Research Mentor}\\
Max Vo, UC Davis \hfill Winter 2025--Present

\begin{itemize}
\tightlist
\item
  Mentored an undergraduate student in developing an interactive Shiny
  app using R to visualize and analyze intra-cluster variance. This app
  will work as a user-friendly interface for the \texttt{ivd} R package,
  enabling researchers to explore and analyze complex datasets
  interactively, enhancing accessibility for non-technical users.
\end{itemize}

\section{SUBSTANTIAL ACTIVITIES DURING GRADUATE
SCHOOL}\label{substantial-activities-during-graduate-school}

\hrule

\begin{itemize}
\tightlist
\item
  Led a break-out room for international students in the virtual event
  \emph{Pathways to Graduate School in Quantitative Methods} held in the
  Fall of 2023. This event seeks to instruct prospective students
  interested in pursuing Quantitative methods graduate degrees.
\end{itemize}

\section{SKILLS}\label{skills}

\hrule

\emph{Programming}

\begin{itemize}
\tightlist
\item
  R: exploratory data analysis, data cleaning, data visualization,
  statistical data analyses, programmatic reporting (R Markdown/
  Quarto), web scraping, dashboards, packages development, Shiny apps
  development.
\item
  Python: exploratory data analysis, data cleaning, data visualization.
\end{itemize}

\emph{Tools}

\begin{itemize}
\tightlist
\item
  Statistical softwares: R, Stan, SPSS, JASP, JAMOVI
\item
  Reporting: RMarkdown, Quarto, \LaTeX
\item
  Version control: Git and Github
\item
  Reference management: Zotero
\item
  Data collection and management: REDCap
\end{itemize}

\end{document}